Estimating real-time highstreet footfall from Wi-Fi probe requests

### 2018-06-25 paper/sections/methodology.tex
Everything except main study is complete

The next step after collecting data was to estimate the footfall or pedestrian activity from them. We identified the following major challanges which arise from our collection methodology.
\begin{enumerate}
	\item \textbf{Background noise} - since the extent to which Wi-Fi signals travel differs subject to various factors such as interference and humidity, it is close to impossible to restrict our data collection to a finite area of interest. This can lead to a signicant background noise at certain locations. E.g. a phone shop or a bus stop located next to the study area can increase the number of probe requests received by the sensor.
	\item \textbf{MAC randomisation} - The mobile devices in the past few years have been using randomised 'local' MAC addresses for probe requests to protect the users from being tracked. This makes it impossible to tell if the probe requests are being sent by the same mobile device which is being stationed next to the sensor. This along with the previous problem can further increase the magnitude of error by several fold.
	\item \textbf{Mobile ownership} - Since the rate of mobile ownership can vary widely across geography and demography, we cannot assume that every mobile device translates to one pedestrian footfall. In addition to this, there is a long term overall increase in mobile ownership which may lead to the number of probe requests collected overtime. 
\end{enumerate}
We propose the following methods to tackle each of these challages.

>>>>

The next step after collecting data was to estimate the footfall or pedestrian activity from them. We identified the following major challanges which arise from our collection methodology.
\begin{enumerate}
	\item \textbf{Background noise} - since the extent to which Wi-Fi signals travel differs subject to various factors such as interference and humidity, it is close to impossible to restrict our data collection to a finite area of interest. This can lead to a signicant background noise at certain locations. E.g. a phone shop or a bus stop located next to the study area can increase the number of probe requests received by the sensor.
	\item \textbf{MAC randomisation} - The mobile devices in recent years have been using randomised 'local' MAC addresses for probe requests to protect the users from being tracked. This makes it impossible to tell if the probe requests are being sent by the same mobile device which is being stationed next to the sensor. This along with the previous problem can further increase the magnitude of error by several fold.
	\item \textbf{Mobile ownership} - Since the rate of mobile ownership can vary widely across geography and demography, we cannot assume that every mobile device translates to one pedestrian footfall. In addition to this, there is a long term overall increase in mobile ownership which may lead to the number of probe requests collected overtime. 
\end{enumerate}
We propose the following methods to tackle each of these challages.

### 2018-06-25 paper/sections/pilot_study.tex
Everything except main study is complete

To start we design a small pilot study as a proof of concept with the aim to validate the classification and clustering methodology against the scale and complexity of data collected on a open public area such as a retail highstreet.
We also want to find the algorithm which is best suited for classification of signal strengths.
The data was collected at Oxford Street in London on 20 December 2017 from 12:30 to 13:00 hrs, where Wi-Fi probe requests sets were collected using the Wi-Fi sensor described earlier and pedestrian footfall was manually recorded using the Android app.
Being located at one of the busiest retail locations in the United Kingdom, the WiFi sensor captured approximately 60,000 probe requests during the period, and 3,722 people were recorded manually.

@@

We then move on to assign an unique field identifying the mobile device generating the probe requests.
For the 45\% of the probe requests which were not randoised, we kept the MAC addresses as the unique identifier.
For the rest of the probe requesrt we applied the clustering alogorithm to assign the unique identifier.
With the help of the known stationary device (the mobile device used by the surveyor to record pedestrians manually) and through trail and error, We found the time threshold, $\alpha$ to be 15 seconds and threshold for sequence numbers, $\beta$ to be 60.
Figure \ref{pilot_clustering} shows the clustering of probe requests.
We finally combine both normal and anonymised probe requests, aggregate them based on their unique identifier which further reduced the mean error to -18\%.

\begin{figure}
	\begin{center}
		\includegraphics [width=\linewidth] {images/pilot_clustering.jpeg}
		\caption{Clustering probe requests based on increasing sequence numbers present in them. The dots are individual probe requests and the red lines connect probe requests within the same cluster which are generated by the same mobile device.}
		\label{pilot_clustering}
	\end{center}
\end{figure}

@@

\begin{figure}
	\begin{center}
		\includegraphics [width=0.85\linewidth] {images/pilot_counts_comparision.png}
		\caption{Comparision of counts after filtering with manual counts}
		\label{pilot_comparison}
	\end{center}
\end{figure}

>>>>

To start we designed a small pilot study with the aim to validate the classification and clustering methodology against the scale and complexity of data collected on a open public area such as a retail highstreet.
We also want to find the algorithm which is best suited for classification of signal strengths.
The data was collected at Oxford Street in London on 20 December 2017 from 12:30 to 13:00 hrs, where Wi-Fi probe requests sets were collected using the Wi-Fi sensor described earlier and pedestrian footfall was manually recorded using the Android app.
Being located at one of the busiest retail locations in the United Kingdom, the WiFi sensor captured approximately 60,000 probe requests during the period, and 3,722 people were recorded manually.

\begin{figure}
	\begin{center}
		\includegraphics [width=0.95\linewidth] {images/pilot_clustering.jpeg}
		\caption{Clustering probe requests based on increasing sequence numbers present in them. The dots are individual probe requests and the red lines connect probe requests within the same cluster which are generated by the same mobile device.}
		\label{pilot_clustering}
	\end{center}
\end{figure}

@@

We then move on to assign an unique field identifying the mobile device generating the probe requests.
For the 45\% of the probe requests which were not randomised, we kept the MAC addresses as the unique identifier.
For the rest of the probe requests we applied the clustering algorithm to assign the unique identifier.
With the help of the known stationary device (the mobile device used by the surveyor to record pedestrians manually) and through trail and error, We found the suitable time threshold, $\alpha$ to be 15 seconds and threshold for sequence numbers, $\beta$ to be 60.
Figure \ref{pilot_clustering} shows the clustering of probe requests.
We finally combine both normal and anonymised probe requests, aggregate them based on their unique identifier which further reduced the mean error to -18\%.

\begin{figure}
	\begin{center}
		\includegraphics [width=0.90\linewidth] {images/pilot_comparison.jpeg}
		\caption{Comparision of counts after filtering with manual counts}
		\label{pilot_comparison}
	\end{center}
\end{figure}

@@

### 2018-06-25 paper/sections/previous_work.tex
Everything except main study is complete

There have been numerous attempts at using Wi-Fi to measure the volume and movement of people in the built environment for various applications \citep{zarim2006,sap2015,reki2007}.
Though most research obtains feasible and favorable results, in recent years, one of the major challenges faced in such attempts has been the MAC address randomisation process.
This process aims to protect the users’ privacy by anonymising the only globally identifiable portion of the probe requests, which results in a set of probe requests generated by the same device with different random MAC addresses \citep{green2008}.
There have been various successful attempts by researchers to breaking this randomisation process in order to extract real MAC addresses, \citep{martin2017} but this usually results in serious risk of infringement of the privacy of the users of the mobile devices.
There is a clear gap in the research for exploring methodologies which enable us to estimate the number of unique mobile devices from a set of anonymised probe requests, without the need to reveal their original MAC addresses.

>>>>

Though WiFi is a `location-less' technology, there are reliable methods to triangulate the location of Wi-Fi enabled mobile devices by the known locations of APs and the signal strength reported by them from the mobile devices \citep{he2003range, moore2004robust, lamarca2005place}.
This can overcome the usual shortcoming of GPS, which struggles for precision and accuracy in indoor and densely built environments \citep{zarim2006,kawaguchi2009wifi, xi2010locating}. 
Utilising this, we can easily and quickly estimate trajectories of the mobile devices just using the WiFi communication the device has with multiple known APs \citep{Sorensenlocation} which can be used similar to the GPS trajectories to understand individual travel patterns (Kim, 2006;\citep{reki2007,Sap2015}, crowd behaviour \citep{abedi2013bluetooth,mowafi2013tracking}, vehicular \citep{lu2010vehicle} and pedestrian movement \citep{xu2013pedestrian,fukuzaki2014pedestrian,wang2016gait}.
Such data can also be used in transportation planning and management to estimate travel time \citep{musa2011wiflow} and real time traffic monitoring \citep{abbott2013empirical}.

Using techniques demonstrated by \cite{franklin2006passive} and \cite{pang2007802} and globally unique information present in the probe requests one can also successfully track people across these access points \citep{cunche2014linking}, their trajectories \citep{musa2012tracking}, and the interactions between them \citep{cheng2012inferring,barbera2013signals,cunche2014know} such as predicting which of them are most likely to meet again \citep{cunche2012know}.
Using the semantic information present in these probe requests it even is possible to understand the nature of population at a large scale \citep{di2016mind}.

Though extensive research has been done on this subject with feasible and favorable results, in recent years, one of the major challenges faced in such attempts has been the increasing attempt by mobile phone manufacturers to protect the users’ privacy by anonymising the globally identifiable portion of the probe requests, \citep{green2008}.
There are various methods which have been devised to overcome this anonymisation process such as decomposition of OUIs where detailed device model information is estimated by analysing an already known dataset of OUIs \citep{martin2016decomposition}; Scrambler attack using a small part of the physical layer specification for WiFi \citep{vo2016,bloessl2015scrambler}; and finally, the timing attack where the packet sequence information present in the probe request frame is used \citep{matte2016,cheng2016can}.
A combination of these methodologies has been proven to produce de-anonymised globally unique device information \citep{vanhoef2016, martin2017}. These approaches usually result in serious risk of infringement of the privacy of the users of the mobile devices by revealing their identifiable personal information.

There is a clear gap in the research for exploring methodologies which enable us to estimate the number of unique mobile devices from a set of anonymised probe requests, without the need to reveal their original MAC addresses.
Such technique has various applications in numerous fields such as uncovering the urban wireless landscape \citep{rose2010mapping}, revealing human activity at large scales \citep{qin2013discovering}, estimating pedestrian numbers in crowds \citep{schauer2014estimating,fukuzaki2015statistical} and even counting people in hyper local scales such as queues \citep{wang2013measuring}
With enough infrastructure to collect such information we can aim to generate a real-time census of the city \citep{konto2017}.
It has also been demonstrated by \citep{pinelli2015comparing} through series of experiments on a telecom operator dataset that using such network-driven approach is more advantageous compared to the widely used event-driven approaches.

With this background we set out to device and implement a methodology to reliably estimate human activity such as pedestrian footfall from Wi-Fi probe requests without infringing the privacy of the users involve.
